---
title: "Longfin Inshore Squid (\\protect\\textit{Doryteuthis pealeii}) \\linebreak Snapshot Ecosystem & Socioeconomic Profile (ESP)"
format: 
  pdf:
    margin-left: "0.25in"
    margin-right: "0.25in"
    margin-top: "0in"
    margin-bottom: "0.25in"
    fontsize: "9pt"
    linestretch: 1
    indent: true
    include-in-header: template.tex    
 #   documentclass: "article"
    keep-tex: true
    papersize: letter
params:
  widths: !expr c(0.9, 0.75, 3, 3)
  img1: !expr here::here("images/longfin_life_history_0602.png")
  img2: !expr here::here("images/risk_plot_new.png")
  img3: !expr here::here("images/squibs_data.png")
  img4: !expr here::here("images/noaa_logo.png")
  tbl_file: !expr here::here("docs/table_template.xlsx")
  img_dir: !expr here::here("images")
latex_engine: xelatex
bibliography: references.bib
csl: cambridge-university-press-numeric.csl
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# here::i_am("inst/esp_template/06_docs/report card pg 1.qmd")
#devtools::load_all()
devtools::load_all(here::here("../READ-EDAB-NEesp2"))
``` 

\backgroundsetup{
  scale=1,
  angle=0,
  opacity=0,
  contents={\includegraphics[width=\paperwidth,height=\paperheight]{bg_pg1.jpg}}
 }
\BgThispage 

\vspace{-2.0cm}
\section{Spring 2026} 

::::: {layout="[[0.575, 0.025, 0.4]] "}

:::: {}

:::{}
\raggedright
\section{Key Findings from the Life History Working Group}

\subsection{Lifespan and aging}
Some literature sources estimate growth to be 1 statolith ring/day, and literature review supports a lifespan of less than 1 year. Participants at the longfin squid summit estimated a maximum age of 15 months. 2024 statolith aging from SQUIBS indicates maximum ages of 7 months for females and 8.6 months for males (right) from squid caught in the fishery.

\vspace{0.25cm}

\subsection{Maturity (from SQUIBS)}
In 2024, most stage 4 squid caught in summer with very little mature squid caught the rest of the year. Highest numbers of stage 1 squid were caught in the second half of 2024. Of 912 squid assessed, the dominant maturity stage in females increases from fall to spring. The highest percentage of mature male squid were caught in spring and summer. No stage 4 females and very few stage 1 males were caught.

\vspace{0.25cm}

\subsection{Migration and movement dynamics}
In November/December, longfin migrate from the inshore shelf to deep, warmer slope waters along the shelf edge to overwinter. By May/June of the next year, they migrate back to shallow coastal waters from the MAB to southern New England to spawn [@Macy_Brodziak_2001]. Recent work hypothesizes the possibility of a winter cohort that hatches south of Cape Hatteras and migrates onto the Northeast shelf [@richardson_2026]. Fishery observations describe a spatial gradient of 1-6 cm mantle length (ML) squid from waters south of Hatteras through southern New England, with the smallest squid detected further south. The Gulf Stream and warm core rings may facilitate the recruitment transport of juvenile squid, but potential for inputs to the population from the South and offshore are difficult to quantify.

\vspace{0.25cm}

\subsection{Reproductive dynamics}
Spawning peaks inshore from late spring to early summer in the Mid-Atlantic and southern New England [@Hatfield_Cadrin_2002] [@Macy_Brodziak_2001] with hatching in late summer [@jacobson_2005]. Consideration of the hypothesis of a winter cohort  spawning south of Hatteras indicates the presence of multiple cohorts of longfin squid, with some outside of the traditional Northeast shelf stock area, and provides evidence of year-round spawning in the stock.

\vspace{0.25cm}

\subsection{Natural mortality}
Although natural mortality is expected to be age-dependent, lack of accurate age data makes further study difficult. Using the equation derived by Hamel and Cope [@Hamel_Cope_2022], natural mortality for longfin squid can range from 0.36 (max. age = 15 mo.) to 0.675 (max. age = 8 mo.). Intraspecific predation impacts natural mortality, but there is no available data to quantify the amount of mortality this causes.
:::

:::{}
![](`r params$img1`)
:::

::::

::::{}
\hfill
::::

::::{}
:::{}
\vspace{0.50cm}
\section{Age Frequency from SQUIBS}
![](`r params$img3`)
:::

:::{}
\section{Key Points from the Mid-Atlantic Risk Assessment}

\raggedright
<!-- maybe this is being centered because it's the only section without a header? -->
The 2025 Mid-Atlantic EAFM Risk Assessment Update [@mafmc_midatlantic_2025] determined that there are moderate-high risks of :

- Potential and observed distribution shifts of longfin squid
- Not achieving optimum yield due to interactions with non-MAFMC managed species
- Regulatory complexity negatively impacting optimum yield due to occasional recent changes in regulations and moderate (3-4) recreational regulation differences across states
- Not minimizing bycatch and discards to the extent practicable due to regular, managed discards and incidental catch and moderate discard mortality

Risk elements are aspects that may threaten achieving the biological, economic, or social objectives that the MAFMC desires from a fishery; risk to achieving optimal yield. 

Longfin squid did not score in the “high” risk category for any risk elements in 2025.

:::

:::{}
\vspace{0.25cm}

![](`r params$img2`)
:::
::::

:::::

\newpage

\backgroundsetup{
  contents={\includegraphics[width=\paperwidth,height=\paperheight]{bg_pg2.jpg}}
  }
\BgThispage 
<!-- not sure why this doesn't work -->

\newgeometry{top=0.25in, left=0.25in, right=0.25in, bottom=0.25in}

```{r}
data <- readxl::read_excel(params$tbl_file) |>
  janitor::clean_names() |>
  dplyr::mutate(figure =  paste0(params$img_dir, "/", time_series)) |>
  dplyr::select(-time_series)
```

::: {}
```{r}
NEesp2::rpt_card_table(data, widths = params$widths) %>% 
  flextable::font(i = 1:5, j = 1:3, fontname='Times New Roman')
```

\section{Data Gaps/Uncertainty}

\raggedright
- Bottom temperature data comes from GLORYS [@jeanmichel_copernicus_2021], a modeled re-analysis product that incorporates insitu data.

- The Gulf Stream Index indicator is a yearly value and may not be indicative of changes in oceanographic processes on a smaller time scale. 

- While literature generally supports a life span around 1 year, various studies and data indicate a maximum age anywhere between 6 to 15 months. A large range of potential maximum ages creates uncertainty around life history processes, spawning timing, and natural mortality.

- A hypothesis has been proposed regarding longfin squid spawning South of Cape Hatteras in the winter months. While survey data in the 1970s and 80s indicate larval squid in this region that are transported north into the Mid-Atlantic Bight, there is a lack of definitive data to prove this hypothesis. If one assumes this hypothesis is true, further questions arise concerning maximum age and inputs to the population from South Atlantic Bight spawning. 

- Lack of a definitive maximum age of longfin squid affects calculations of natural mortality. Effects of cannibalism on the population are unknown at this time.

\vspace{1.0cm}

\centering\normalsize

We welcome your observations! Please contact [northeast.ecosystem.highlights@noaa.gov](northeast.ecosystem.highlights@noaa.gov) with any on-the-water insights or changes observed in the black sea bass fishery and [nefsc.esp.leads@noaa.gov](nefsc.esp.leads@noaa.gov) with questions or comments on the information presented in this report.\newline

The code used to create this report can be viewed online: [github.com/NEFSC/READ-EDAB-longfinESP](github.com/NEFSC/READ-EDAB/longfinESP) 


\newpage
\newgeometry{top=1in, left=1in, right=1in, bottom=1in}

# References
::: {#refs}
:::